---
title: "Exercise 1"
author: "Carilli"
date: '`r format(lubridate::today(), "%B %d, %Y")`'
output: 
  html_document:
    theme: cerulean
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = FALSE, message = FALSE,
                      options(digits = 3, scipen = 999))
library(tidyverse)
```

```{r 1data}
height_wage <- 
  here::here("data","HeightWage_MenWomenUS_HW.csv") %>% 
  read_csv()
```

### (a)

```{r 1a}
height_wage %>% 
  mutate(sex = factor(male, labels = c("female", "male"))) %>% 
  lm(wage96 ~ height85 + height81 + sex, .) %>% 
  jtools::summ()
```

### (b)

```{r 1b}
height_wage %>% 
  mutate(female = 1 - male,
         male = factor(male, labels = c("female", "male")),
         female = factor(female, labels = c("female", "male"))) %>% 
  lm(wage96 ~ height85 + height81 + male + female, .) %>% 
  jtools::summ()
```

### (c)

```{r 1C}
height_wage %>% 
  mutate(sex = factor(male, labels = c("female", "male"))) %>% 
  filter(sex == "male") %>% 
  lm(wage96 ~ height85 + height81, .) %>% 
  jtools::summ()

height_wage %>% 
  mutate(sex = factor(male, labels = c("female", "male"))) %>% 
  filter(sex == "female") %>% 
  lm(wage96 ~ height85 + height81, .) %>% 
  jtools::summ()
```

### (d)

```{r 1d}
height_wage %>% 
  mutate(sex = factor(male, labels = c("female", "male"))) %>% 
  lm(wage96 ~ height85 + height81 + sex * height81 + sex * height85 - sex, .)  %>% 
  jtools::summ()
```

### (e)

```{r 1e}
height_wage %>% 
  mutate(sex = factor(male, labels = c("female", "male"))) %>% 
  lm(wage96 ~ height85 + height81 + sex * height81 + sex * height85, .)  %>% 
  jtools::summ()
```

### (f)

```{r 1f}
height_wage %>% 
  mutate(sex = factor(male, labels = c("female", "male"))) %>% 
  lm(wage96 ~ height81 * sex + norest96 + norcen96 + south96, .) %>% 
  jtools::summ()

height_wage %>% 
  mutate(sex = factor(male, labels = c("female", "male"))) %>% 
  lm(wage96 ~ height81 * sex + norest96 + norcen96 + west96, .) %>% 
  jtools::summ()
```


#### Tidy Data Principles

Columns are variables rows are observations.  We have race spread across three columns and region spread across 4 columns.  Tidy data suggests that both region and race be one variable (column) each and be factors not numeric.


If the following returns false, there is an other category for race, if it returns true, black, white, Hispanic exhaust the racial categories.

```{r}
height_wage %>% 
  summarize(sum(black,white,hispanic)) == nrow(height_wage)
```

Let's combine black, white, Hispanic into race.

```{r}
height_wage %>% 
  mutate(white = ifelse(white == 1,  "white", ""),
         black = ifelse(black == 1, "black", ""),
         hispanic = ifelse(hispanic == 1, "hispanic", ""),
         race = factor(str_c(white, black, hispanic))) %>% 
  select(-black, -hispanic, -white)
```

We can do the same for region

```{r}
height_wage %>% 
  mutate(norest96 = ifelse(norest96 == 1, "norest96", ""),
         norcen96 = ifelse(norcen96 == 1, "norcen96", ""),
         south96 = ifelse(south96 == 1, "south96", ""),
         west96 = ifelse(west96 == 1, "west96", ""),
         region = factor(str_c(norest96, norcen96, south96, west96))) %>% 
  select(region, norest96, norcen96, south96, west96)
```

Finally, let's convert male into a factor sex.

```{r}
height_wage %>% 
  mutate(sex = factor(male, labels = c("male", "female")))
```

Now we can combine these pipelines into one to create a tidy data frame.

```{r}
height_wage_tidy <- 
height_wage %>% 
  mutate(white = ifelse(white == 1,  "white", ""),
         black = ifelse(black == 1, "black", ""),
         hispanic = ifelse(hispanic == 1, "hispanic", ""),
         race = factor(str_c(white, black, hispanic)),
         norest96 = ifelse(norest96 == 1, "norest96", ""),
         norcen96 = ifelse(norcen96 == 1, "norcen96", ""),
         south96 = ifelse(south96 == 1, "south96", ""),
         west96 = ifelse(west96 == 1, "west96", ""),
         region = factor(str_c(norest96, norcen96, south96, west96)),
         sex = factor(male, labels = c("male", "female"))) %>% 
  select(-white, -black, -hispanic, -norest96, -norcen96, -south96, -west96, -male)
```

Part f with tidy data.

```{r}
height_wage_tidy %>% 
  lm(wage96 ~ height81 * sex + region, .) %>% 
  jtools::summ()

height_wage_tidy %>% 
  lm(wage96 ~ height81 * sex + relevel(region, "west96"), .) %>% 
  jtools::summ()
```

